\textbf{Are regime changes always bad economics? Evidence from daily financial data}

### 2019-12-04 doc/paper.tex
Updates in line with reviewer comments

\subsection{Reexamining how markets respond to irregular regime changes}

As noted above, conventional theory and cross-country empirical evidence suggests that political instability depresses financial returns. However, there are reasons to believe such hypotheses may be overly simplistic.

First, political instability may be seen as a positive event if the current regime is regarded as strongly anti-business or anti-global. Regime changes that attempt to overthrow anti-business autocrats may be expected to result in positive abnormal returns as investors see little risk of a replacement government ``worse'' than the status quo. \citet{alesina1996political} take the possibility of a new government adopting better economic policies seriously, but argue that the negative effects of uncertainty dominate the positive effects of coups staged by pro growth factions.

An active debate exists on the subject of ``good coups.'' \citet{alesina1996income} argue that ``unconstitional'' regime changes such as coups are worse for economic growth than other types of political instability, while   \citet{londregan1990poverty} find that coups have no impact on economic growth and hypothesize that this may be because there is a bimodal distribution of coups, some of which enhance growth and others which restrict it. However, most research suggests that ``good coups'' are not the norm \citep{derpanopoulos2015coups, powell2011global, thyne2016coup, varol2011democratic}. But while ``good coups'' may be the minority, some evidence suggests that they have become more frequent \citep{marinov2014coups} and could have positive economic effects. \citet{meyersson2016political} finds that coup attempts that occur in democratic regimes are associated with negative GDP growth, while those that occur in autocratic regimes are associated with positive GDP growth, and \citet{girardi2018institution} show that Salvador Allende's socialist government was associated with large declines in stock prices while the coup that replaced him boosted them.

Previous research therefore suggests that on average coups should not be expected to lead to increased economic development and positive market returns, but that positive returns associated with ``good coups'' could occasionally be present. We provide the first systematic empirical evidence in support of this theory as our event study approach also allows us examine ``good coups''---coups that lead to democratization or economic liberalization---separately from others. Unlike \citet{girardi2018institution}, we are not limited to a single case study, and can therefore provide both a cross-country and within-case analysis.

Second, not all political instability is equivalent--- resignations, assassinations, and coups are all examples of unexpected regime changes. However, resignations typically occur when an unpopular leader steps down and may signal the coming of a more effective leader. Assassinations are not always related to the effectiveness of a leader, can occur seemingly at random, and the successor to the assassinated leader may be unclear. Coups can occur in the name of democracy or autocracy, and in general signal relative weakness in the current political system. These differences are often overlooked in past empirical strategies, which have proxied for instability in terms of the number of coups \citep{londregan1990poverty, alesina1996political}, assassinations or revolutions \citep{barro1991economic}, or have combined various metrics into single indices \citep{alesina1996income, venieris1986income, gupta1990economics}.\footnote{Note that these studies vary by metrics included in the indices, method of aggregation, and outcome of interest.} In another cross-country regression study, \citet{jong2009measurement} disaggregates instability into four main metrics, one of which is regime changes, but does not subset below this metric. By contrast---recognizing the different degrees of instability that these events may reflect---we estimate effects separately for coups, assassinations, resignations, and protests. Our event study approach affords us this freedom without encountering the power issues that would arise in a regression setting.

Different types and events of instability may therefore have disparate effects on markets or growth. Taken together, the theory and research above suggests that coups should on average lead to negative returns, but sometimes lead to positive returns when the coup's instigators are clearly more pro market than the regime they replace. Assassinations should have a neutral or negative impact as they always increase uncertainty, but institutional responses to assassinations vary by country. By contrast, resignations may be viewed positively on average, as they typically signal the departure of an ineffective leader.

@@

The results in the figure are particularly striking given the discrepancy between the ARs on event days 0 and 1 and all other days. Consistent with our earlier findings that coups tend to have pre and post-event CARs that are statistically indistinguishable from zero, the only days on which the 2002 Venezuelan coup d'etat attempt ARs are statistically different from zero is on event days 0 and 1 after the coup attempt. The almost 0\% 10-day CAR preceding the coup makes this an ideal case as it implies that investors were unaware of the coup plot. This implies that both events were unexpected by investors, increasing our confidence that the abnormal returns capture the true effect of the Chavez to Carmona and Carmona to Chavez regime changes on stock returns. We see the same story 10 years prior, with pre and post-event CARs statistically indistinguishable from zero on all days except for the day of the Chavez coup attempt, this time demonstrating the market's negative evaluation of a Pérez to Chavez regime change.

@@

These results are consistent with our hypothesis that the nature of the political event and its expected impact on policy matters. While the mean effect of assassinations is negative, it is smaller in magnitude than for coups. Unlike a coup, an assassination may not necessarily be expected to cause immediate change in economic policy, particularly in the presence of an institutionalized line of succession. As such, we would expect CARs to be negative due to increased instability and uncertainty, but smaller in magnitude to a coup or resignation due to greater expectations of policy inertia.

@@

These results are consistent with the idea that the expected impact of an atypical political event on economic policy is highly important to the directionality of returns. The positive event day abnormal return following resignations is not surprising since resignations typically occur because of poor performance and/or loss of authority. Leaders were often ousted following corruption charges, allegations of fraud, financial crises, and/or political violence.

@@

@@

Column (1) suggests that public protests have no effect on stock returns. However, this occurs because some political movements increase stock prices while others decrease them. As shown in column (2), the absolute value of stock returns are over 2\% higher during public protests. These estimates would be biased if protest dates are correlated with higher world or regional stock market indices. To address this potential confounder, column (3) controls for returns on the S\&P/IFC Emerging Markets Investable Composite Stock Index. The coefficient on \textit{Protest} barely changes and the absolute value of returns are still about 2\% higher during public protests. Finally, column (4) shows that stock volatility is more than 1 percentage point higher during political movements.\footnote{Volatility estimation methodology is described in detail in \autoref{subsec: Volatility}.}

>>>>

\singlespacing
\section{Reexamining how markets respond to irregular regime changes}
\doublespacing

Conventional theory and cross-country empirical evidence suggests that political instability depresses financial returns. However, there are reasons to believe such hypotheses may be overly simplistic.

First, political instability may be seen as a positive event if the current regime is regarded as anti-business or anti-global. Regime changes that attempt to overthrow anti-business autocrats may be expected to result in positive abnormal returns as investors see little risk of a replacement government ``worse'' than the status quo. \citet{alesina1996political} take the possibility of a new government adopting better economic policies seriously, but argue that the negative effects of uncertainty dominate the positive effects of coups staged by pro growth factions.

An active debate exists on the subject of ``good coups''---i.e. coups that lead to democratization or economic liberalization. \citet{alesina1996income} argue that ``unconstitional'' regime changes such as coups are worse for economic growth than other types of political instability, while \citet{londregan1990poverty} find that coups have no impact on economic growth and hypothesize that this may be because there is a bimodal distribution of coups, some of which enhance growth and others which restrict it. However, most research suggests that ``good coups'' are not the norm \citep{derpanopoulos2015coups, powell2011global, thyne2016coup, varol2011democratic}. But while ``good coups'' may be the minority, some evidence suggests that they have become more frequent \citep{marinov2014coups} and could have positive economic effects. \citet{meyersson2016political} finds that coup attempts that occur in democratic regimes are associated with negative GDP growth, while those that occur in autocratic regimes are associated with positive GDP growth, and \citet{girardi2018institution} show that Salvador Allende's socialist government was associated with large declines in stock prices while the coup that replaced him boosted them.

Previous research therefore suggests that on average coups should not be expected to lead to increased economic development and positive market returns, but that positive returns associated with ``good coups'' could occasionally be present. We shed light on this theory as our event study approach also allows us to examine ``good coups''---coups that lead to democratization or economic liberalization---separately from others. As we are not limited to a single case study, we can provide both a cross-country and within-case analysis.

Second, not all political instability is equivalent--- resignations, assassinations, and coups are all examples of unexpected regime changes. However, resignations typically occur when an unpopular leader steps down and may signal the coming of a more effective leader. Assassinations are not always related to the effectiveness of a leader, can occur seemingly at random, and the successor to the assassinated leader may be unclear. Coups can occur in the name of democracy or autocracy, and in general signal relative weakness in the current political system. These differences are often overlooked in past empirical strategies, which have proxied for instability in terms of the number of coups \citep{londregan1990poverty, alesina1996political}, assassinations or revolutions \citep{barro1991economic}, or have combined various metrics into single indices \citep{alesina1996income, venieris1986income, gupta1990economics}.\footnote{Note that these studies vary by metrics included in the indices, method of aggregation, and outcome variable of interest.} In another cross-country regression study, \citet{jong2009measurement} disaggregates instability into four main metrics, one of which is regime changes, but does not subset below this metric. By contrast---recognizing the different degrees of instability that these events may reflect---we estimate effects separately for coups, assassinations, resignations, and protests. Our event study approach affords us this freedom without encountering the power issues that would arise in a regression setting.

Different types and events of instability may therefore have disparate effects on markets or growth. Taken together, the theory and research above suggests that coups should on average lead to negative returns, but sometimes lead to positive returns when the coup's instigators are clearly more pro market than the regime they replace. Assassinations should have a neutral or negative impact as they increase uncertainty, but institutional responses to assassinations vary by country. By contrast, resignations may be viewed positively on average, as they typically signal the departure of an ineffective leader.

@@

The results in \autoref{fig:AR-Ven} are particularly striking given the discrepancy between the ARs on event days 0 and 1 and all other days. Consistent with our earlier findings that coups tend to have pre and post-event CARs that are statistically indistinguishable from zero, the only days on which the 2002 Venezuelan coup d'etat attempt ARs are statistically different from zero is on event days 0 and 1 after the coup attempt. The almost 0\% 10-day CAR preceding the coup makes this an ideal case as it implies that investors were unaware of the coup plot. This implies that both events were unexpected by investors, increasing our confidence that the abnormal returns capture the true effect of the Chavez to Carmona and Carmona to Chavez regime changes on stock returns. We see the same story 10 years prior, with pre and post-event CARs statistically indistinguishable from zero on all days except for the day of the Chavez coup attempt, this time demonstrating the market's negative evaluation of a Pérez to Chavez regime change.

@@

These results are consistent with our hypothesis that the nature of the political event and its expected impact on policy matters, and that assassinations should have a negative effect as they occur seemingly at random and increase uncertainty. While the mean effect of assassinations is negative, it is smaller in magnitude than for coups. Unlike a coup, an assassination may not necessarily be expected to cause immediate change in economic policy, particularly in the presence of an institutionalized line of succession. As such, we would expect CARs to be negative due to increased instability and uncertainty, but smaller in magnitude to a coup or resignation due to greater expectations of policy inertia.

@@

These results are again consistent with our hypothesis that different events will have disparate effects, and that resignations may lead to positive returns. The positive event day abnormal return following resignations is not surprising as resignations typically occur because of poor performance and/or loss of authority. Among our sample of events, leaders were ousted following loss in conflict/war, anti-authoritarian protest, corruption scandals, Supreme Court ruling against unconstitutional actions, contested elections, and abuse of power.

@@

By contrast, the largest negative event in our sample (-3\%) is the 1993 resignation of President Ghulam Ishaq Khan and Prime Minister Nawaz Sharif in Pakistan. The resignations occurred after months of political infighting when the army demanded the President and Prime Minister resign and call for new elections. An interim prime minister was installed, but uncertainty about Pakistan's political and economic future remained high prior to the next round of elections.

@@

Column (1) suggests that public protests have no effect on stock returns. However, this occurs because some political movements increase stock prices while others decrease them. As shown in column (2), the absolute value of stock returns are approximately 1.5\% higher during public protests. These estimates would be biased if protest dates are correlated with higher world or regional stock market indices. To address this potential confounder, column (3) controls for returns on the S\&P/IFC Emerging Markets Investable Composite Stock Index. The coefficient on \textit{Protest} barely changes and the absolute value of returns are still about 1.5\% higher during public protests. Finally, column (4) shows that stock volatility is approximately 1 percentage point higher during political movements.\footnote{Volatility estimation methodology is described in detail in \autoref{subsec: Volatility}.}
